---
title: "Mid Term"
author: "Saleh"
date: "November 1, 2019"
output: pdf_document
---


```{r}
#install.packages("ISLR")
#install.packages("dplyr")
#install.packages("caret")
#install.packages("tidyverse")
#install.packages("factoextra")
#install.packages("tidyverse")
#install.packages("flexclust")


library(ISLR)
library(dplyr)
library(caret)

library(tidyverse)  
library(factoextra) 

library(flexclust)



Universty <- read.csv("Universities.csv")

set.seed(123)
# Or this method:

#Remove all records with missing measurements from the dataset.
Uni <- na.omit(Universty)

# Using only the continuous measurements
U_continuous <- Uni[,-c(1,2,3)]
```


```{r}
# B
# we will scale the data first
df <- scale(U)
distance <- get_dist(df)
fviz_dist(distance) # plot

# K-Means algorithm to find out the clusters 

k4 <- kmeans(df, centers = 4, nstart = 25) # k = 4, number of restarts = 25

k4$centers 

k4$size # Number of universities in each cluster


k4$cluster[120] 


fviz_cluster(k4, data = df) # Visualize the output(plot)


# Elbow Method
fviz_nbclust(df, kmeans, method = "wss")


# Compute k-means clustering with k = 4
set.seed(123)
k3 <- kmeans(df, centers =3, nstart = 25)
print(k3)

fviz_cluster(k3, data = df) # Cluster Plot

# Silhouette Method


fviz_nbclust(df, kmeans, method = "silhouette")

```

```{r}
#Creating the cluster index for 3 clusters
set.seed(123)
k3_median = kcca(U, k=3, kccaFamily("kmedians"))
clusters_index_3 <- predict(k3_median)
```

```{r}
#install.packages(stats)
library(stats)
#install.packages("HAC")
library(HAC)
Uni_scale <- scale(U, center=T, scale= T)
U.dis <- dist(Uni_scale)
cah.ward <- hclust(U.dis, method = "ward.D2")
plot(cah.ward)

# partition in 3 groups
groupes.cah <- cutree(cah.ward,k=3)
```


```{r}
#Function for calculating summary statistics -y cluster membership variable
stat.comp <-function(x,y){ #number of clusters
K <- length(unique(y))
#nb. Of instances
n <- length(x)
#overall mean
m <- mean(x)
#total sum of squares
TSS <- sum((x-m)^2)
#size of clusters
nk <- table(y)
#conditional mean
mk <- tapply(x,y,mean)
#between (explained) sum of squares
BSS <- sum(nk * (mk-m)^2)
#collect in a vector the means and the proportion of variance explained
result <- c(mk,100.0*BSS/TSS)
#set a name to the values
names(result) <- c(paste("G",1:K),"% epl.")
#return the results
return(result)
}
#applying the function to the original variables of the dataset
#and not to the standardized variables

print(sapply(U,stat.comp ,y=groupes.cah))
```

```{r}
#Merging the clusters to the original data frame
set.seed(123)
clusters <- data.frame(clusters_index_3)
Universities <- cbind(Universities, clusters)
Universities$room_board_fees <- Universities$room + Universities$board + Universities$add..fees + Universities$estim..book.costs + Universities$estim..personal..
```


```{r}
#  4
#All
set.seed(123)
Summary_cont <- Universities %>% 
    group_by(clusters_index_3) %>%
    summarise( Acceptance_rate = sum(X..appl..accepted)/ sum(X..appli..rec.d), Avg_out_state_tuition=mean(out.of.state.tuition), Avg_int_state_tuition=mean(in.state.tuition), room_board_fees=mean(room_board_fees), mean_PHD_fac=mean(X..fac..w.PHD), mean_stud_fac_ratio=mean(stud..fac..ratio), mean_grad_rate=mean(Graduation.rate), priv_count = sum(Public..1...Private..2. == 2), pub_count = sum(Public..1...Private..2. == 1))
Summary_cont

#Private
Summary_cont_priv <- Universities %>% filter(Public..1...Private..2. == 2) %>%
    group_by( clusters_index_3 ) %>%
    summarise( Acceptance_rate = sum(X..appl..accepted)/ sum(X..appli..rec.d), Avg_out_state_tuition=mean(out.of.state.tuition), Avg_int_state_tuition=mean(in.state.tuition), room_board_fees=mean(room_board_fees), mean_PHD_fac=mean(X..fac..w.PHD), mean_stud_fac_ratio=mean(stud..fac..ratio), mean_grad_rate=mean(Graduation.rate))
Summary_cont_priv


#Public
Summary_cont_pub <- Universities %>% filter(Public..1...Private..2. == 1) %>%
    group_by( clusters_index_3 ) %>%
    summarise( Acceptance_rate = sum(X..appl..accepted)/ sum(X..appli..rec.d), Avg_out_state_tuition=mean(out.of.state.tuition), Avg_int_state_tuition=mean(in.state.tuition), room_board_fees=mean(room_board_fees), mean_PHD_fac=mean(X..fac..w.PHD), mean_stud_fac_ratio=mean(stud..fac..ratio), mean_grad_rate=mean(Graduation.rate))
Summary_cont_pub
```
D - There is a relationship between the category variable. cluster 3 is the only cluster with a majority public schools. This explains the low average tution rates.

E - Some additional information that could help explain the data would be the state of the school, or the operating budget of the university, or the amount of academic endowments of the university.There is traditionally a lot of differences between public and private schools in this regard and could explain why some of the public schools fell into a different cluster.

```{r}
#f. Isolating the data to Tufts University cluster 2 had the lowest distance

library(dplyr)
library(stats)
norm_Tufts <- scale(Universities1[, c(4, 5, 6, 7, 8, 9, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20)])
Tufts <- filter(Universities1, College.Name == "Tufts University")

#Euclidean Distance for Each Clusters it is clear that cluster 1 has the lowest distance from Tufts
#Cluster 1
dist(rbind(Tufts[, -c(1, 2, 3, 10)], k3$centers[1,]))
#Cluster 2
dist(rbind(Tufts[, -c(1, 2, 3, 10)], k3$centers[2,]))
#Cluster 3
dist(rbind(Tufts[, -c(1, 2, 3, 10)], k3$centers[3,]))
```

```{r}
#Impute NAs with average from cluster 2
cluster_2 <- filter(Universities, clusters_index_3 == 2)
avg_cluster_2 <- mean(cluster_2[,c(10)])
Tufts[, c(10)] <- avg_cluster_2
Tufts[, c(10)]
```

